### 2024-12-05 paper structure/notes.tex
Paper structure

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Ideas on paper structure}
\maketitle

\section{\textbf{Abstract}}
Things to cite:
\begin{itemize}
  \item Design of \textbf{dynamic} ETM for Discrete Time \textbf{non-linear} systems stabilized by neural network controller
  \item Events' triggering logic based on the use of local sector conditions for the non-linear activation functions, saturation
  \item Development of sufficient LMI conditions for ETM parameters' design together with the computation of an inner approximation of the region of attraction
  \item Convex optimization procedures to achieve different objectives in the final system behavior. (e.g. size of ROA, number of events, etc.)
  \item \textbf{Sara dice che l'Abstract si fa per ulitmo e ha ragione}
\end{itemize}

\section{Introduction}

\section{Problem statement}
\subsection{System description}
\subsection{Controller description}
\subsection{Event-triggering Mechanism description}

\section{}

% \pagebreak
% \printbibliography

\end{document}