### 2024-12-06 paper structure/theory.tex
Schemes created

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Theory draft}
\maketitle

\subsection*{\textbf{Notation}}
$\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$, and the set of real matrices of dimension $n \times m$, respectively. For any matrix $A$, $A\tr$ is its transpose. For any square matrix $A$ the operator $He\left\{ A \right\} = A + A\tr$ is defined. $\texttt{diag}(A_1, A_2)$ is a block-diagonal matrix with block diagonal matrices $A_1$ and $A_2$. For a partitioned matrix the symbol $\star$ stands for symmetric blocks.

\section*{Problem statement}

\section*{\textbf{Lemma}}
Considering Finsler's lemma in the context of our ETM, with respect to system (\dots) and layer $i \in \left[ 0, \dots, l \right]$, and assuming that (\dots) holds $\forall \xi \in \left[ \underbar{\xi},  \bar{\xi} \right]$, if $\exists X_i \in \RR^{\left( n_x + 2 \cdot n_{\phi_i} \right) \times\left( n_x + 2 \cdot n_{\phi_i} \right)}, N_{1}^{i} \in \RR^{n_{\phi} \times n_x}, N_2^i \in \RR^{n_{\phi} \times n_{\phi}}, N_3^i \in \RR^{n_{\phi} \times n_{\phi}}$

\end{document}