### 2024-12-09 Paper draft/draft.tex
Update up to ETM desing. TODO Finsler lemma

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\title{Draft of paper}

\author{Marco Sterlini}

\section{Introduction}

\section{Problem Formulation}
We consider the non-linear discrete-time system stabilized by a Neural Network controller represented in figure \ref{fig:first_scheme} described by:
\begin{equation}
  x^{+} = A x + B \text{sat}(\bar{u}) + C \Phi(E x) + D d 
\end{equation}
with $x \in \RR^{n_x}$ the state vector, $\bar{u}$ the raw output of the controller, $\Phi$ the representation of the non-linearity of the system and $d$ a constant disturbance input.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/first_scheme}
    \caption{Feedback system}
    \label{fig:first_scheme}
\end{figure}

The controller is implemented as a Multi Layer Perceptron with $l$ layers, each with $n_{\phi_i}$ neurons, all the activation functions are saturation functions, represented by $\phi_i$. The objective is to design an event-triggering mechanism that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller. The controller $\pi$ is defined by:
\begin{equation}\label{eq:nn-equations}
  \begin{aligned}
  \omega^{0}(k) &= x(k) \\
  \nu^{i}(k) &= W^{i} \omega^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
  \omega^{i}(k) &= \phi_i(\nu^i(k))\\
  \bar{u}(k) &= \phi_l(\nu^l(k))
  \end{aligned} 
\end{equation}
with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi_i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k) \in \RR^{n_{\phi_i}}$ the output. The weights $W^i \in \RR^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \RR^{n_{\phi_i}}$ define the affine operation of each layer. The saturation function application is to be intended element-wise and will be denoted as $\phi_i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron. Adopting the same notation as in \cite{css-paper} we define the controller policy in a condensed form. Denoting the augmented vectors:
\begin{equation*}
  \nu_{\phi} = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega_{\phi} = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi_{\nu_{\phi}} = \bmx{\phi_1(\nu^1) \\ \vdots \\ \phi_l(\nu^l)} 
\end{equation*}
with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega_\phi = \phi(\nu_\phi)$. Finally, conditions \ref{eq:nn-equations} can be rewritten as:
\begin{equation*}
  \bmx{u(k)\\ \nu_\phi(k)} = N \bmx{x(k) \\ \omega_\phi(k) \\ 1} 
\end{equation*}
where
\begin{equation}\label{eq:first-N}
  \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & W^l & b^l \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^{l-1} & \0 & b^{l-1}
    \end{array}
  }
\end{equation}

Similarly as in \cite{css-extended} we implement an event-triggering-mechanism (ETM) in every layer of the controller, we add also a final ETM to the output that will be used to further reduce the total number of events and call to the non-linear functions. The system and the controller are now described by figure \ref{fig:second_scheme}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{img/second_scheme}
    \caption{Feedback system subject to ETM in the controller}
    \label{fig:second_scheme}
\end{figure}

The dynamics of the system are now described by:
\begin{equation}\label{eq:system-dynamics}
  x^{+} = A x + B u + C \Phi(E x) + D d
\end{equation}
with $u$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \ref{eq:first-N} as:
\begin{equation}\label{eq:last-N}
  N = \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & \Id & 0 \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^l & \0 & b^l
    \end{array}
  } = \bmx{
    \begin{array}{c | c | c}
    N_{u x} & N_{u \omega} & N_{u b} \\
    \hline
    N_{\nu x} & N_{\nu \omega} & N_{\nu b}
    \end{array} 
  }
\end{equation}

The system is compliant with the assumptions of \textbf{Lemma 2} in \cite{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:

\begin{equation}
    \begin{aligned}
         R &= (\Id - N_{\nu \omega})^{-1}\\
         R_{\omega} &= N_{u x} + N_{u \omega} R N_{\nu x}\\
         R_b &= N_{u \omega} R N_{\nu b} + N_{u b}
    \end{aligned}
\end{equation}

With $N_{\nu \omega}$ always invertible due to its lower triangular structure. We express:

\begin{equation}
  \bmx{u_* \\ \nu_*} = N \bmx{x_* \\ \omega_* \\ 1}; \ \  
  \begin{aligned}
    \omega_* &= \nu_*\\
    \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
    u_* &= R_{\omega} x_* + R_b
  \end{aligned}
\end{equation}

The equilibrium state $x^*$ needs some further discussion since the expression to retrieve it is no longer explicit due to the non-linearity $\Phi$. We can retrieve the vector using a numerical solver that minimizes the following expression for a given disturbance $\bar{d}$:

\begin{equation}
  (A + B R_{\omega} - \Id) x_* + C \Phi(E x_*) + D \bar{d} + B R_b = 0
\end{equation}

\section{Main Results}
Taking inspiration from \textbf{Proposition 1} and \textbf{Lemma 3} in \cite{css-extended} we design the ETM for the output of each layer of the controller and the final output. The idea is to exploit the sector conditions of the activation functions as ETM conditions.

\subsection{Event-triggering mechanism}
An ETM works by triggering an event that will update the current layer and propagate its value across the network. The triggering function will take as inputs the incremental variables of the state of the system 
$\tilde{x}$, the current state of the layer $\tilde{\nu}$ and the dead-zone vector $\tilde{\psi} = \tilde{\nu} - \tilde{\omega}$. The ETM was defined as:
\begin{equation}
  \omega^{i}(k) = \begin{cases}
    \omega^{i}(k-1) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > 0\\
    \phi_i(\nu^i(k)) & \text{otherwise}
  \end{cases}
\end{equation}

In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form:

\begin{equation}\label{eqn:base_omega}
  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = 
  \tilde{\psi}\tr T^i \left[ G^i \tilde{x} + \tilde{\psi} - \tilde{\nu} \right] = 
  \bmx{\tilde{x}\\ \tilde{\psi}^i \\ \tilde{\nu}^i}\tr
  \bmx{
    0 & 0 & 0\\
    T^i G^i & T^i & -T^i\\
    0 & 0 & 0
  } 
  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}} > 0
\end{equation}

This approach acts as a static ETM triggering an event as soon as the triggering function $f$ is positive. Inspired by the work of \cite{data-driven} we define a dynamic threshold to be implemented in the ETM. The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system reducing the overall number of events and computational burden. The new triggering function is defined as:

\begin{equation}
  \omega^{i}(k) = \begin{cases}
    \omega^{i}(k-1) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta_i\\
    \phi_i(\nu^i(k)) & \text{otherwise}
  \end{cases};
\end{equation}
\begin{equation}
  \eta^+_i = \rho^i \eta_i - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu})\  \forall i \in \left( 1, \dots, l \right) 
\end{equation}
with $0 < \rho^i < 1 \ \forall i \in \left( 1, \dots, l \right)$. Since the previous conditions hold for each layer $i$ the notation can be simplified by denoting with:

\begin{equation}
  R = \text{diag} \left\{ \rho^1, \dots, \rho^l \right\}; \ \Psi = \text{diag}\left\{ f^1, \dots, f^l \right\}; \ \eta = \text{diag} \left\{ \eta^1, \dots, \eta^l \right\}
\end{equation}

The ETM triggering condition can be rewritten as: $\Psi > R \eta$ in an element-wise fashion.

\subsection{Lyapunov conditions}

\subsection{Optimization procedure}

\section{Simulations}

\section{Conclusions}

\printbibliography

\end{document}